%% ===================================================================
\section{Experimental Results}
\label{sec:eval}


\subsection{Polybench/C (30 Kernels)}

\begin{table}[t]
  \caption{Polybench/C results.  WA = with analysis, NA = no analysis.}
  \label{tab:results_1x}
  \centering\small
  \begin{tabular}{lcc}
    \toprule
    \textbf{Metric} & \textbf{WA} & \textbf{NA} \\
    \midrule
    Pass rate              & 30/30 (100\%) & 28/30 (93\%) \\
    Median speedup         & 2.15$\times$  & 1.06$\times$ \\
    Mean speedup           & 3.44$\times$  & 1.52$\times$ \\
    Kernels $>1\times$     & 20/30         & 14/28         \\
    \midrule
    \multicolumn{3}{l}{\emph{Head-to-head (28 common):}} \\
    WA wins ($>$10\% faster) & \multicolumn{2}{c}{12} \\
    NA wins ($>$10\% faster) & \multicolumn{2}{c}{6 (3 race conditions)}  \\
    Ties (within 10\%)       & \multicolumn{2}{c}{10}  \\
    \bottomrule
  \end{tabular}
\end{table}

Table~\ref{tab:results_1x} summarizes results at standard
(1$\times$) problem sizes.  All 30 kernels pass with analysis
guidance; without it, two stencil kernels fail due to
parallelizing dimensions with carried dependences.
The 2.15$\times$ median speedup is limited by Polybench's
small defaults ($N$\,=\,60--120), where GPU launch overhead
dominates.  Compute-bound kernels show strong speedups
(\texttt{heat\_3d}: 12.36$\times$, \texttt{deriche}:
7.44$\times$, \texttt{covariance}: 7.40$\times$).

In the head-to-head comparison (Table~\ref{tab:results_1x}), 6 of 28 common kernels appear faster without analysis guidance. However, 3 of these are \emph{false wins}: the unguided LLM parallelizes loop dimensions that carry data dependencies, producing code that runs faster but computes incorrect results. For example, in \texttt{lu}, the NA code launches one GPU thread per row (\texttt{grid=(N,)}), but row~$i$ depends on the result of row~$k$ ($k<i$). Without sequential ordering, threads race to read unfinished rows, causing output differences up to $10^{33}$ across repeated runs. With analysis guidance, \sysname correctly identifies this dependence and enforces sequential row processing, sacrificing parallelism for correctness. The remaining 3 NA wins are within measurement noise and reflect nondeterministic LLM code quality rather than a systematic advantage of unguided generation.

\paragraph{Scaling.}

\begin{table}[t]
  \caption{Polybench scaling: 1$\times$ to 8$\times$ problem sizes
    ($N$: 60$\to$480, matrices: 120$\to$960).
    ``ST'' = single-threaded C; ``OMP'' = 80-thread OpenMP.}
  \label{tab:results_8x}
  \centering\small
  \begin{tabular}{lccc}
    \toprule
    \textbf{Metric} & \textbf{1$\times$} & \textbf{8$\times$ (ST)} & \textbf{8$\times$ (OMP)} \\
    \midrule
    Pass rate          & 30/30        & 30/30         & 30/30        \\
    Benchmarked        & 30           & 25$^*$        & 25$^*$       \\
    Median speedup     & 2.15$\times$ & 30.73$\times$ & 8.57$\times$ \\
    Kernels $>1\times$ & 20/30        & 21/25         & 20/25        \\
    \bottomrule
    \multicolumn{4}{l}{\scriptsize $^*$5 sequential kernels timeout during benchmark.}
  \end{tabular}
\end{table}

At 8$\times$ sizes (Table~\ref{tab:results_8x}), median speedup
rises to 30.73$\times$ over single-threaded C.  Against 80-thread
OpenMP, the GPU still achieves 8.57$\times$ median, confirming
genuine advantage beyond CPU parallelism.  Five sequential kernels
time out during benchmark (\texttt{grid=(1,)}).

\paragraph{Guidance Impact.}

\begin{table}[t]
  \caption{Impact of analysis-triggered guidance categories.
    ``Before'' and ``After'' are speedups at 1$\times$ sizes
    except where noted.}
  \label{tab:guidance_fixes}
  \centering\small
  \begin{tabular}{llcc}
    \toprule
    \textbf{Guidance} & \textbf{Kernel} & \textbf{Before} & \textbf{After} \\
    \midrule
    Multi-phase split      & correlation & 0.19$\times$ & 2.50$\times$ \\
    WAR sequential         & nussinov    & FAIL          & PASS         \\
    Triangular grid        & lud$^\dagger$ & 0.14$\times$ & 13.18$\times$ \\
    Fused reductions       & bicg        & 1.04$\times$ & 3.51$\times$ \\
    Stencil barriers       & jacobi\_1d  & 0.13$\times$ & 1.78$\times$ \\
    IIR phase fusion       & deriche     & 0.28$\times$ & 7.19$\times$ \\
    Tridiag.\ sweep        & adi         & 0.24$\times$ & 2.27$\times$ \\
    \bottomrule
    \multicolumn{4}{l}{\scriptsize $^\dagger$Rodinia benchmark.}
  \end{tabular}
\end{table}

Table~\ref{tab:guidance_fixes} shows the impact of individual
guidance categories, each triggered by analysis properties
(e.g., cross-phase dependences, triangular patterns).
The largest gains come from multi-phase decomposition
(correlation: 13$\times$) and triangular grid
(lud: 94$\times$).

\paragraph{Limitations.}
Ten kernels achieve $<$1$\times$ at standard sizes: five are
inherently sequential (loop-carried dependences limit parallelism),
and two (\texttt{jacobi\_2d}, \texttt{fdtd\_2d}) suffer from
kernel launch overhead at small $N$---at 8$\times$ sizes they reach
36.82$\times$ and 15.43$\times$.


\subsection{TSVC (151 Kernels)}
\label{sec:tsvc}

\begin{table}[t]
  \caption{TSVC results at three data sizes.  ``ST'' = single-threaded
    C (\texttt{-O2}), ``OMP'' = OpenMP C (\texttt{-O2 -fopenmp},
    80~threads). Pass rate: 150/151 (99.3\%).}
  \label{tab:tsvc}
  \centering\small
  \begin{tabular}{lcc}
    \toprule
    & \textbf{32\,MB} & \textbf{3200\,MB} \\
    \midrule
    Per-array size     & 32\,MB & 3.2\,GB \\
    Parallelizable$^*$ & 27/150 & 15/150  \\
    \midrule
    Tri/ST median      & 67.1$\times$ & 73.2$\times$ \\
    Tri/OMP CPU median & 10.7$\times$ & 14.2$\times$ \\
    Tri/OMP GPU median & 1.02$\times$ & ---           \\
    \bottomrule
    \multicolumn{3}{l}{\scriptsize $^*$Serial/2D kernels excluded at large sizes.}
  \end{tabular}
\end{table}
We apply \sysname to 151 TSVC~\cite{callahan1988tsvc} kernels
covering linear traversals, reductions, indirect addressing,
and loop-carried dependences.
Table~\ref{tab:tsvc} shows results at realistic data sizes.
\sysname passes 150/151 (99.3\%), 120 on first attempt.
At 32\,MB, LLM-generated Triton achieves 67$\times$ (ST)
and 10.7$\times$ (80-thread OMP CPU).  Crucially, Triton
achieves \textbf{1.02$\times$ vs.\ OpenMP GPU offloading}
(\texttt{nvc~-mp=gpu}) on the \emph{same} GPU, showing the
LLM produces code matching a production compiler.


\subsection{Application Kernels}
\label{sec:realworld}

\begin{table}[t]
  \caption{Application kernels: Triton (WA) vs.\ OMP~GPU offloading
    (\texttt{nvc}) on the same GPU, and WA vs.\ NA ablation (vs.\ ST).
    ``---'' = failed.}
  \label{tab:realworld}
  \centering\small
  \begin{tabular}{llccc}
    \toprule
    \textbf{Source} & \textbf{Kernel} & \textbf{Tri/OMP\,GPU} & \textbf{WA/ST} & \textbf{NA/ST} \\
    \midrule
    Rodinia & SRAD       & 24.9$\times$  & 24.9$\times$ & 3.5$\times$ \\
    Rodinia & hotspot    & 4.3$\times$   & 1.1$\times$  & 1.2$\times$ \\
    Rodinia & gaussian   & 249$\times$   & 3.4$\times$  & 3.6$\times$ \\
    miniMD  & LJ force   & 6.8$\times$   & 25.8$\times$ & 23.0$\times$ \\
    miniFE  & SpMV       & 3.1$\times$   & 9.0$\times$  & 7.3$\times$ \\
    \midrule
    Rodinia & lud        & ---           & 13.2$\times$ & 15.5$\times$ \\
    Rodinia & lavaMD     & ---           & 1.5$\times$  & 0.6$\times$ \\
    Rodinia & pathfinder & ---           & 0.1$\times$  & --- \\
    \bottomrule
  \end{tabular}
\end{table}

To evaluate generalization, we apply \sysname to 6 Rodinia
kernels~\cite{che2009rodinia} and 2 ECP proxy
kernels~\cite{heroux2009mantevo} (Table~\ref{tab:realworld}).
All 8 pass correctness with analysis guidance (6 on first attempt).

\paragraph{Triton vs.\ OMP GPU (same hardware).}
On 5 kernels with OMP~GPU baselines, LLM-generated Triton
is 3--249$\times$ faster than \texttt{nvc}'s OpenMP offloading.
The advantage comes from Triton's block-level model: the LLM
fuses phases, uses shared memory tiling, and avoids
per-timestep target-region launch overhead that penalizes
\texttt{nvc} on iterative kernels (gaussian: 511 timesteps
$\times$ 3~launches = 1533~offloads).  SRAD's multi-phase
stencil benefits similarly (24.9$\times$).

\paragraph{WA vs.\ NA (analysis impact).}
Analysis guidance is most impactful on multi-phase kernels:
SRAD improves from 3.5$\times$ (NA) to 24.9$\times$ (WA)
via phase decomposition; lavaMD improves from 0.56$\times$ to
1.47$\times$ as analysis identifies the parallelizable dimension
(NA fails on 4/5 attempts).
